Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-06-24 survey/survey.tex
one sentence on graphical
just comments

@@

@@

\begin{figure}[tb]
Let $c_1 : t_1 \leftrightarrow t_2$,  $c_2 : t_2 \leftrightarrow t_3$, and $c_3 : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {c_1 \odot (c_2 \odot c_3) \Leftrightarrow (c_1 \odot c_2) \odot c_3}
\\
  {(c_1 \oplus (c_2 \oplus c_3)) \odot \assoclp \Leftrightarrow \assoclp \odot ((c_1 \oplus c_2) \oplus c_3)}
\\
  {(c_1 \otimes (c_2 \otimes c_3)) \odot \assoclt \Leftrightarrow \assoclt \odot ((c_1 \otimes c_2) \otimes c_3)}
\\
  {((c_1 \oplus c_2) \oplus c_3) \odot \assocrp \Leftrightarrow \assocrp \odot (c_1 \oplus (c_2 \oplus c_3))}
\\
  {((c_1 \otimes c_2) \otimes c_3) \odot \assocrt \Leftrightarrow \assocrt \odot (c_1 \otimes (c_2 \otimes c_3))}
\\
  {\assocrp \odot \assocrp \Leftrightarrow ((\assocrp \oplus \idc) \odot \assocrp) \odot (\idc \oplus \assocrp)}
\\
  {\assocrt \odot \assocrt \Leftrightarrow ((\assocrt \otimes \idc) \odot \assocrt) \odot (\idc \otimes \assocrt)}
\end{array}\]
\caption{\label{figj}Signatures of level-2 $\Pi$-combinators: associativity}
\end{figure}

@@

\begin{figure}[tb]
Let $a : t_1 \leftrightarrow t_2$, $b : t_3 \leftrightarrow t_4$, and $c : t_5 \leftrightarrow t_6$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {((a \oplus b) \otimes c) \odot \dist \Leftrightarrow \dist \odot ((a \otimes c) \oplus (b \otimes c))}
\\
  {(a \otimes (b \oplus c)) \odot \distl \Leftrightarrow \distl \odot ((a \otimes b) \oplus (a \otimes c))}
\\
  {((a \otimes c) \oplus (b \otimes c)) \odot \factor \Leftrightarrow \factor \odot ((a \oplus b) \otimes c)}
\\
  {((a \otimes b) \oplus (a \otimes c)) \odot \factorl \Leftrightarrow \factorl \odot (a \otimes (b \oplus c))}
\end{array}\]
\caption{\label{figi}Signatures of level-2 $\Pi$-combinators: distributivity and factoring}
\end{figure}

@@

\begin{figure}[tb]
Let $c, c_1, c_2, c_3 : t_1 \leftrightarrow t_2$ and $c', c'' : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\idc \odot \, c \Leftrightarrow c}
\quad
  {c \, \odot \idc \, \Leftrightarrow c}
\quad
  {c\,\, \odot\,!\, c \Leftrightarrow \idc}
\quad
  {!\, c \odot c \Leftrightarrow \idc}
\\
  {c \Leftrightarrow c}
\quad
\Rule{}
  {c_1 \Leftrightarrow c_2 \quad c_2 \Leftrightarrow c_3}
  {c_1 \Leftrightarrow c_3}
  {}
\quad
\Rule{}
  {c_1 \Leftrightarrow c' \quad c_2 \Leftrightarrow c''}
  {c_1 \odot c_2 \Leftrightarrow c' \odot c''}
  {}
\end{array}\]
\caption{\label{figh}Signatures of level-2 $\Pi$-combinators: identity and composition}
\end{figure}

@@

\begin{figure}[tb]
Let $c_0 : 0 \leftrightarrow 0$, $c_1 : 1 \leftrightarrow 1$, and $c : t_1 \leftrightarrow t_2$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\identlp \odot c \Leftrightarrow (c_0 \oplus c) \odot \identlp}
\qquad
  {\identrp \odot (c_0 \oplus c) \Leftrightarrow c \odot \identrp}
\\
  {\identlsp \odot c \Leftrightarrow (c \oplus c_0) \odot \identlsp}
\qquad
  {\identrsp \odot (c \oplus c_0) \Leftrightarrow c \odot \identrsp}
\\
  {\identlt \odot c \Leftrightarrow (c_1 \otimes c) \odot \identlt}
\qquad
  {\identrt \odot (c_1 \otimes c) \Leftrightarrow c \odot \identrp}
\\
  {\identlst \odot c \Leftrightarrow (c \otimes c_1) \odot \identlst}
\qquad
  {\identrst \odot (c \otimes c_1) \Leftrightarrow c \odot \identrst}
\\
  {\identlt \Leftrightarrow \distl \odot (\identlt \oplus \identlt)}
\\
\identlp \Leftrightarrow \swapp \odot \identlsp
\qquad
\identlt \Leftrightarrow \swapt \odot \identlst
\end{array}\]
\caption{\label{figg}Signatures of level-2 $\Pi$-combinators: unit}
\end{figure}

@@

\begin{figure}[tb]
Let $c_1 : t_1 \leftrightarrow t_2$ and $c_2 : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\swapp \odot (c_1 \oplus c_2) \Leftrightarrow (c_2 \oplus c_1) \odot \swapp}
\quad
  {\swapt \odot (c_1 \otimes c_2) \Leftrightarrow (c_2 \otimes c_1) \odot \swapt}
\\
  {(\assocrp \odot \swapp) \odot \assocrp \Leftrightarrow ((\swapp \oplus \idc) \odot \assocrp) \odot (\idc \oplus \swapp)}
\\
  {(\assoclp \odot \swapp) \odot \assoclp \Leftrightarrow ((\idc \oplus \swapp) \odot \assoclp) \odot (\swapp \oplus \idc)}
\\
  {(\assocrt \odot \swapt) \odot \assocrt \Leftrightarrow ((\swapt \otimes \idc) \odot \assocrt) \odot (\idc \otimes \swapt)}
\\
  {(\assoclt \odot \swapt) \odot \assoclt \Leftrightarrow ((\idc \otimes \swapt) \odot \assoclt) \odot (\swapt \otimes \idc)}
\end{array}\]
\caption{\label{figf}Signatures of level-2 $\Pi$-combinators: commutativity and associativity}
\end{figure}

@@

\begin{figure}[tb]
Let $c_1 : t_1 \leftrightarrow t_2$, $c_2 : t_3 \leftrightarrow t_4$, $c_3 : t_1 \leftrightarrow t_2$, and $c_4 : t_3 \leftrightarrow t_4$. \\
Let $a_1 : t_5 \leftrightarrow t_1$,  $a_2 : t_6 \leftrightarrow t_2$, $a_3 : t_1 \leftrightarrow t_3$, and $a_4 : t_2 \leftrightarrow t_4$.
\[\def\arraystretch{1.3}
\begin{array}{c}
\Rule{}
  {c_1 \Leftrightarrow c_3 \quad c_2 \Leftrightarrow c_4}
  {c_1 \oplus c_2 \Leftrightarrow c_3 \oplus c_4}
  {}
\qquad
\Rule{}
  {c_1 \Leftrightarrow c_3 \quad c_2 \Leftrightarrow c_4}
  {c_1 \otimes c_2 \Leftrightarrow c_3 \otimes c_4}
  {}
\\
  {\idc \oplus \, \idc \, \Leftrightarrow \idc}
\qquad
  {\idc \otimes \, \idc \, \Leftrightarrow \idc}
\\
  {(a_1 \odot a_3) \oplus (a_2 \odot a_4) \Leftrightarrow (a_1 \oplus a_2) \odot (a_3 \oplus a_4)}
\\
  {(a_1 \odot a_3) \otimes (a_2 \odot a_4) \Leftrightarrow (a_1 \otimes a_2) \odot (a_3 \otimes a_4)}
\end{array}\]
\caption{\label{fige}Signatures of level-2 $\Pi$-combinators: functors}
\end{figure}

@@

\begin{figure}[tb]
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\identlsp \oplus \idc ~\Leftrightarrow~ \assocrp \odot (\idc \oplus \, \identlp)}
\\
  {\identlst \otimes \idc ~\Leftrightarrow~ \assocrt \odot (\idc \otimes \, \identlt)}
\end{array}\]
\caption{\label{figd}Signatures of level-2 $\Pi$-combinators: unit and associativity}
\end{figure}

@@

\begin{figure}[tb]
Let $c : t_1 \leftrightarrow t_2$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {(c \otimes \idc) \odot \absorbl \Leftrightarrow \absorbl \odot \idc}
\quad
  {(\idc \, \otimes c) \odot \absorbr \Leftrightarrow \absorbr \odot \idc}
\\
  {\idc \odot \, \factorzl \Leftrightarrow \factorzl \odot (\idc \otimes c)}
\quad
  {\idc \odot \, \factorzr \Leftrightarrow \factorzr \odot (c \otimes \idc)}
\\
  {\absorbr \Leftrightarrow \absorbl}
\\
  {\absorbr \Leftrightarrow (\distl \odot (\absorbr \oplus \absorbr)) \odot \identlp}
\\
  {\identlst \Leftrightarrow \absorbr}
\qquad
  {\absorbl \Leftrightarrow \swapt \odot \absorbr}
\\
  {\absorbr \Leftrightarrow (\assoclt \odot (\absorbr \otimes \idc)) \odot \absorbr}
\\
  {(\idc \otimes \absorbr) \odot \absorbl \Leftrightarrow (\assoclt \odot (\absorbl \otimes \idc)) \odot \absorbr}
\\
  {\idc \otimes \, \identlp \Leftrightarrow (\distl \odot (\absorbl \oplus \idc)) \odot \identlp}
\end{array}\]
\caption{\label{figc}Signatures of level-2 $\Pi$-combinators: zero}
\end{figure}

@@

\begin{figure}[tb]
\[\def\arraystretch{1.3}
\begin{array}{c}
  {((\assoclp \otimes \idc) \odot \dist) \odot (\dist \oplus \idc) \Leftrightarrow (\dist \odot (\idc \oplus \dist)) \odot \assoclp}
\\
  {\assoclt \odot \distl \Leftrightarrow ((\idc \otimes \distl) \odot \distl) \odot (\assoclt \oplus \assoclt)}
\end{array}\]
\vspace{ -0.5em}
\[\def\arraystretch{1.3}
\begin{array}{rcl}
  (\distl \odot (\dist \oplus \dist)) \odot \assoclp &\Leftrightarrow&
   \dist \odot (\distl \oplus \distl) \odot \assoclp ~\odot \\
&& (\assocrp \oplus \idc) ~\odot \\
&& ((\idc \oplus \swapp) \oplus \idc) ~\odot \\
&&      (\assoclp \oplus \idc)
\end{array}\]
\caption{\label{figb}Signatures of level-2 $\Pi$-combinators: associativity and distributivity}
\end{figure}

@@

\begin{figure}[tb]
\[\def\arraystretch{1.3}
\begin{array}{rcl}
  (\idc \otimes \swapp) \odot \distl &\Leftrightarrow& \distl \odot \swapp
\\
  \dist \odot (\swapt \oplus \swapt) &\Leftrightarrow & \swapt \odot \distl
\end{array}\]
\caption{\label{figa}Signatures of level-2 $\Pi$-combinators: commutativity and distributivity}
\end{figure}

@@

>>>>

%% Random idea:
%% for a product monoidal this is wave-style
%% for a sum monoidal this is particle-style
%% when we have both we do have a wave-particle duality

@@

Other than distributity, $\Pi$ combinators simply allow us to shuffle
the components of a stack around as well as shuffle the positions of
components in the plane.

@@

\begin{figure}[t]
Let $c_1 : t_1 \leftrightarrow t_2$,  $c_2 : t_2 \leftrightarrow t_3$, and $c_3 : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {c_1 \odot (c_2 \odot c_3) \Leftrightarrow (c_1 \odot c_2) \odot c_3}
\\
  {(c_1 \oplus (c_2 \oplus c_3)) \odot \assoclp \Leftrightarrow \assoclp \odot ((c_1 \oplus c_2) \oplus c_3)}
\\
  {(c_1 \otimes (c_2 \otimes c_3)) \odot \assoclt \Leftrightarrow \assoclt \odot ((c_1 \otimes c_2) \otimes c_3)}
\\
  {((c_1 \oplus c_2) \oplus c_3) \odot \assocrp \Leftrightarrow \assocrp \odot (c_1 \oplus (c_2 \oplus c_3))}
\\
  {((c_1 \otimes c_2) \otimes c_3) \odot \assocrt \Leftrightarrow \assocrt \odot (c_1 \otimes (c_2 \otimes c_3))}
\\
  {\assocrp \odot \assocrp \Leftrightarrow ((\assocrp \oplus \idc) \odot \assocrp) \odot (\idc \oplus \assocrp)}
\\
  {\assocrt \odot \assocrt \Leftrightarrow ((\assocrt \otimes \idc) \odot \assocrt) \odot (\idc \otimes \assocrt)}
\end{array}\]
\caption{\label{figj}Signatures of level-2 $\Pi$-combinators: associativity}
\end{figure}

@@

\begin{figure}[t]
Let $a : t_1 \leftrightarrow t_2$, $b : t_3 \leftrightarrow t_4$, and $c : t_5 \leftrightarrow t_6$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {((a \oplus b) \otimes c) \odot \dist \Leftrightarrow \dist \odot ((a \otimes c) \oplus (b \otimes c))}
\\
  {(a \otimes (b \oplus c)) \odot \distl \Leftrightarrow \distl \odot ((a \otimes b) \oplus (a \otimes c))}
\\
  {((a \otimes c) \oplus (b \otimes c)) \odot \factor \Leftrightarrow \factor \odot ((a \oplus b) \otimes c)}
\\
  {((a \otimes b) \oplus (a \otimes c)) \odot \factorl \Leftrightarrow \factorl \odot (a \otimes (b \oplus c))}
\end{array}\]
\caption{\label{figi}Signatures of level-2 $\Pi$-combinators: distributivity and factoring}
\end{figure}

@@

\begin{figure}[t]
Let $c, c_1, c_2, c_3 : t_1 \leftrightarrow t_2$ and $c', c'' : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\idc \odot \, c \Leftrightarrow c}
\quad
  {c \, \odot \idc \, \Leftrightarrow c}
\quad
  {c\,\, \odot\,!\, c \Leftrightarrow \idc}
\quad
  {!\, c \odot c \Leftrightarrow \idc}
\\
  {c \Leftrightarrow c}
\quad
\Rule{}
  {c_1 \Leftrightarrow c_2 \quad c_2 \Leftrightarrow c_3}
  {c_1 \Leftrightarrow c_3}
  {}
\quad
\Rule{}
  {c_1 \Leftrightarrow c' \quad c_2 \Leftrightarrow c''}
  {c_1 \odot c_2 \Leftrightarrow c' \odot c''}
  {}
\end{array}\]
\caption{\label{figh}Signatures of level-2 $\Pi$-combinators: identity and composition}
\end{figure}

@@

\begin{figure}[t]
Let $c_0 : 0 \leftrightarrow 0$, $c_1 : 1 \leftrightarrow 1$, and $c : t_1 \leftrightarrow t_2$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\identlp \odot c \Leftrightarrow (c_0 \oplus c) \odot \identlp}
\qquad
  {\identrp \odot (c_0 \oplus c) \Leftrightarrow c \odot \identrp}
\\
  {\identlsp \odot c \Leftrightarrow (c \oplus c_0) \odot \identlsp}
\qquad
  {\identrsp \odot (c \oplus c_0) \Leftrightarrow c \odot \identrsp}
\\
  {\identlt \odot c \Leftrightarrow (c_1 \otimes c) \odot \identlt}
\qquad
  {\identrt \odot (c_1 \otimes c) \Leftrightarrow c \odot \identrp}
\\
  {\identlst \odot c \Leftrightarrow (c \otimes c_1) \odot \identlst}
\qquad
  {\identrst \odot (c \otimes c_1) \Leftrightarrow c \odot \identrst}
\\
  {\identlt \Leftrightarrow \distl \odot (\identlt \oplus \identlt)}
\\
\identlp \Leftrightarrow \swapp \odot \identlsp
\qquad
\identlt \Leftrightarrow \swapt \odot \identlst
\end{array}\]
\caption{\label{figg}Signatures of level-2 $\Pi$-combinators: unit}
\end{figure}

@@

\begin{figure}[t]
Let $c_1 : t_1 \leftrightarrow t_2$ and $c_2 : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\swapp \odot (c_1 \oplus c_2) \Leftrightarrow (c_2 \oplus c_1) \odot \swapp}
\quad
  {\swapt \odot (c_1 \otimes c_2) \Leftrightarrow (c_2 \otimes c_1) \odot \swapt}
\\
  {(\assocrp \odot \swapp) \odot \assocrp \Leftrightarrow ((\swapp \oplus \idc) \odot \assocrp) \odot (\idc \oplus \swapp)}
\\
  {(\assoclp \odot \swapp) \odot \assoclp \Leftrightarrow ((\idc \oplus \swapp) \odot \assoclp) \odot (\swapp \oplus \idc)}
\\
  {(\assocrt \odot \swapt) \odot \assocrt \Leftrightarrow ((\swapt \otimes \idc) \odot \assocrt) \odot (\idc \otimes \swapt)}
\\
  {(\assoclt \odot \swapt) \odot \assoclt \Leftrightarrow ((\idc \otimes \swapt) \odot \assoclt) \odot (\swapt \otimes \idc)}
\end{array}\]
\caption{\label{figf}Signatures of level-2 $\Pi$-combinators: commutativity and associativity}
\end{figure}

@@

\begin{figure}[t]
Let $c_1 : t_1 \leftrightarrow t_2$, $c_2 : t_3 \leftrightarrow t_4$, $c_3 : t_1 \leftrightarrow t_2$, and $c_4 : t_3 \leftrightarrow t_4$. \\
Let $a_1 : t_5 \leftrightarrow t_1$,  $a_2 : t_6 \leftrightarrow t_2$, $a_3 : t_1 \leftrightarrow t_3$, and $a_4 : t_2 \leftrightarrow t_4$.
\[\def\arraystretch{1.3}
\begin{array}{c}
\Rule{}
  {c_1 \Leftrightarrow c_3 \quad c_2 \Leftrightarrow c_4}
  {c_1 \oplus c_2 \Leftrightarrow c_3 \oplus c_4}
  {}
\qquad
\Rule{}
  {c_1 \Leftrightarrow c_3 \quad c_2 \Leftrightarrow c_4}
  {c_1 \otimes c_2 \Leftrightarrow c_3 \otimes c_4}
  {}
\\
  {\idc \oplus \, \idc \, \Leftrightarrow \idc}
\qquad
  {\idc \otimes \, \idc \, \Leftrightarrow \idc}
\\
  {(a_1 \odot a_3) \oplus (a_2 \odot a_4) \Leftrightarrow (a_1 \oplus a_2) \odot (a_3 \oplus a_4)}
\\
  {(a_1 \odot a_3) \otimes (a_2 \odot a_4) \Leftrightarrow (a_1 \otimes a_2) \odot (a_3 \otimes a_4)}
\end{array}\]
\caption{\label{fige}Signatures of level-2 $\Pi$-combinators: functors}
\end{figure}

@@

\begin{figure}[t]
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\identlsp \oplus \idc ~\Leftrightarrow~ \assocrp \odot (\idc \oplus \, \identlp)}
\\
  {\identlst \otimes \idc ~\Leftrightarrow~ \assocrt \odot (\idc \otimes \, \identlt)}
\end{array}\]
\caption{\label{figd}Signatures of level-2 $\Pi$-combinators: unit and associativity}
\end{figure}

@@

\begin{figure}[t]
Let $c : t_1 \leftrightarrow t_2$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {(c \otimes \idc) \odot \absorbl \Leftrightarrow \absorbl \odot \idc}
\quad
  {(\idc \, \otimes c) \odot \absorbr \Leftrightarrow \absorbr \odot \idc}
\\
  {\idc \odot \, \factorzl \Leftrightarrow \factorzl \odot (\idc \otimes c)}
\quad
  {\idc \odot \, \factorzr \Leftrightarrow \factorzr \odot (c \otimes \idc)}
\\
  {\absorbr \Leftrightarrow \absorbl}
\\
  {\absorbr \Leftrightarrow (\distl \odot (\absorbr \oplus \absorbr)) \odot \identlp}
\\
  {\identlst \Leftrightarrow \absorbr}
\qquad
  {\absorbl \Leftrightarrow \swapt \odot \absorbr}
\\
  {\absorbr \Leftrightarrow (\assoclt \odot (\absorbr \otimes \idc)) \odot \absorbr}
\\
  {(\idc \otimes \absorbr) \odot \absorbl \Leftrightarrow (\assoclt \odot (\absorbl \otimes \idc)) \odot \absorbr}
\\
  {\idc \otimes \, \identlp \Leftrightarrow (\distl \odot (\absorbl \oplus \idc)) \odot \identlp}
\end{array}\]
\caption{\label{figc}Signatures of level-2 $\Pi$-combinators: zero}
\end{figure}

@@

\begin{figure}[t]
\[\def\arraystretch{1.3}
\begin{array}{c}
  {((\assoclp \otimes \idc) \odot \dist) \odot (\dist \oplus \idc) \Leftrightarrow (\dist \odot (\idc \oplus \dist)) \odot \assoclp}
\\
  {\assoclt \odot \distl \Leftrightarrow ((\idc \otimes \distl) \odot \distl) \odot (\assoclt \oplus \assoclt)}
\end{array}\]
\vspace{ -0.5em}
\[\def\arraystretch{1.3}
\begin{array}{rcl}
  (\distl \odot (\dist \oplus \dist)) \odot \assoclp &\Leftrightarrow&
   \dist \odot (\distl \oplus \distl) \odot \assoclp ~\odot \\
&& (\assocrp \oplus \idc) ~\odot \\
&& ((\idc \oplus \swapp) \oplus \idc) ~\odot \\
&&      (\assoclp \oplus \idc)
\end{array}\]
\caption{\label{figb}Signatures of level-2 $\Pi$-combinators: associativity and distributivity}
\end{figure}

@@

\begin{figure}[t]
\[\def\arraystretch{1.3}
\begin{array}{rcl}
  (\idc \otimes \swapp) \odot \distl &\Leftrightarrow& \distl \odot \swapp
\\
  \dist \odot (\swapt \oplus \swapt) &\Leftrightarrow & \swapt \odot \distl
\end{array}\]
\caption{\label{figa}Signatures of level-2 $\Pi$-combinators: commutativity and distributivity}
\end{figure}

@@

Idea: thinking in the 3D way; we have collections of towers of
blocks. Every 1-combinator involves touching some blocks, swapping
them around, etc. Two 1-combinators are equivalent if they touch the
same blocks the same number of times. ??
